%&pdflatex
\documentclass[11pt]{article}

\usepackage{main-macros}

\usepackage{newpxmath}
\usepackage{newpxtext}
\usepackage[margin=0.75in]{geometry}


\title{8202 HW4}
\author{Devon Tuma}
\date{Spring 2020}

\begin{document}
\maketitle

\problem{13.5.7}
\begin{proof}
  Let $K$ be an imperfect field of characteristic $p$, so $K^p \ne K$.
  We know that $K^p \subseteq K$, so then there exists $a \in K$ with $a \not \in K^p$, i.e. $k^p \ne a$ for all $k \in K$. Then we argue that $f(x) = x^{p^n} - a$ is an irreducible inseparable polynomial over $K$.
  We have that the formal derivative of this polynomial is $D_x f(x) = p^n x^{p^n - 1} \equiv 0$, and so any root of $f$ is a root of $D_x f$ as well. Hence $f$ is inseparable.

  Then to show it is irreducible, note that if $\alpha$ is a root of $f(x) = x^{p^n} - a$ then $\alpha^{p^n} = a$, and we get $(x - \alpha)^{p^n} = x^{p^n} - a$.
  Hence any factor of $f(x)$ must be of the form $(x - \alpha)^{q}$ for some $1 \le q < p^n$, and furthermore $q$ is of the form $p^k$ for some $k < n$ since the field has characteristic $p$.
  But then the constant term of $(x - \alpha)^{p^k}$ is $\alpha^{p^k}$, which can't be in $K$ or else $a = ((\alpha^{p^k})^{p^{n - k - 1}})^p \in K^p$, which contradicts the definition of $a$.
  Therefore none of the possible factors of $f$ are defined over $K$, and so we conclude $f$ is irreducible. 

  Then finally since $f$ is irreducible and inseparable we have that the extension $K[x]/(f(x))$ is an inseparable finite extension of $K$, as desired.
\end{proof}


\problem{13.5.8}
\begin{proof}
  We proceed by induction on $\deg(f)$.
  When $\deg(f) = 0$ then $f(x) = a_0$ for some $a_0$, and so since $\F{p}$ is perfect we get $f(x)^p = a_0^p = a_0 = f(x^p)$.
  Then for the inductive case, assume the equality holds for polynomials $g$ such that $\deg(g) < n$, and let $f(x)$ have $\deg(f) = n$.
  Then $f$ is of the form $f(x) = a_n x^n + ... + a_0$, and so letting $g(x) = a_{n-1}x^{n-1} + ... + a_0$ we get:
  \begin{align}
    f(x)^p & = (a_n x^n + ... + a_0)^p \\
           & = ((a_n x^n) + (g(x))^p \\
           & = (a_n x^n)^p + g(x)^p \\
           & = a_n^p x^{np} + g(x^p) \\
           & = a_n (x^p)^n + g(x^p) 
             = f(x^p) 
  \end{align}
  Where line $3$ follows from the fact that $(a + b)^p = a^p + b^p$ in a finite field, line $4$ follows from the inductive hypothesis, and line $5$ follows from the fact that $\F{p}$ is perfect.
  Therefore we conclude by induction that $f(x)^p = f(x^p)$ for all $f \in \F_p[x]$ as desired.
\end{proof}


\problem{13.5.9}
\begin{proof}
  We have by definition of binomial coefficients that $\binom{m}{k}$ is the coefficient of $a^k b^{m-k}$ in $(a + b)^m$.
  So taking $b = 1$, $a = x$, $m = pn$, and $k = pi$ we get that $\binom{pn}{pi}$ is the coefficient of $x^{pi} 1^{pn - pi} = x^{pi}$ in $(1 + x)^{pn}$. 
  Then by the previous exercise, we have that $(1 + x)^{pn} = ((1 + x)^p)^n = (1 + x^p)^n$ in $\F_p$, so the coefficient of $(x^p)^i = x^{pi}$ in $(1 + x^p)^n$ is $\binom{pn}{pi}$. 
  But also by the binomial theorem we have that the coefficient of $(x^p)^i$ in $(1 + x^p)^n$ is $\binom{n}{i}$.
  Therefore $\binom{pn}{pi} = \binom{n}{i}$ in $\F_p$ and we conclude $\equivmod{\binom{pn}{pi}}{\binom{n}{i}}{p}$
\end{proof}


\problem{13.6.3}
\begin{proof}
  Assume that $\F$ is a field that contains the $n$th roots of unity for some odd $n$.
  Then let $\zeta$ be a $2n$th root of unity, so $\zeta$ is a root of $f(x) = x^{2n} - 1$, and we need to show $\zeta \in \F$.
  But then note that $f(x) = x^{2n} - 1 = (x^n - 1)(x^n + 1) = -(x^n - 1)((-x)^n - 1)$, so $\zeta$ is either an $n$th root of unity or its negative is.
  But we assumed $\F$ contains all $n$th roots of unity, and fields are closed under inverses so $\F$ also contains the negatives of the $n$th roots of unity, so $\F$ contains all roots of $f$.
  Then since $\zeta$ was a root of $f(x)$ we get that $\zeta \in \F$, and we conclude that $\F$ contains all $2n$th roots of unity as desired.
\end{proof}


\problem{13.6.4}
\begin{proof}
  Note that the $n$th root of unity are the roots of $f(x) = x^n - 1 = x^{p^km} - 1$.
  But then in characteristic $p$ we have that $f(x) = x^{p^km} - 1 = (x^m - 1)^{p^k}$, so the distinct roots of $f(x)$ are the distinct roots of $x^m - 1$, so the distinct $n$th roots of unity are the distinct roots of $x^m - 1$.
  Since $x^m - 1$ has at most $m$ distinct roots, we see that the number of $n$th roots of unity is at most $m$.

  Then to show the number of roots is at least $m$, it suffices to show that $x^m - 1$ and its derivative are relatively prime.
  But note $\frac{d}{dx}(x^m - 1) = mx^{m-1} \ne 0$, where we know it isn't zero since it is relatively prime to the characteristic.
  But then all the roots of $mx^{m-1}$ are $0$, and $0$ is not a root of $x^m - 1$, so $x^m - 1$ polynomial is indeed separable.
  Since all roots of $x^m - 1$ are distinct, we get that the number of $n$th roots of unity is at lest $m$, and finally we conclude that it is exactly $m$ as desired.
\end{proof}


\problem{13.6.5}
\begin{proof}
  Let $\K/\Q$ be a finite extension and let $[\K:\Q] = k$, and assume for contradiction that $\K$ contains infinitely many roots of unity.
  Then consider the cyclotomic polynomials $\Phi_n(x)$.
  Since $\K$ contains infinitely many roots of unity, it contains a root of $\Phi_n$ for arbitrarily large $n$.
  In particular it will contain a root for a $\Phi_m$ with $m$ satisfying $\varphi(m) > k$ (such an $m$ exists since $n$ was arbitrarily large and $\phi$ is strictly increasing while $k$ is fixed).
  But we know that $\deg(\Phi_m) = \varphi(m)$, and by construction $\varphi(m) > k$, so $\K$ contains a root of the polynomail $\Phi_m$ which has degree larger than its extension degree.
  Therefore $\K$ contains the extension generated by this root of $\Phi_m$, which is an extension of degree $\phi(m)$ by the irreducibility of $\Phi_m$.
  But this is a contradiction since extension degree is both multiplicative and integral, so we conclude that $\K$ only contains finitely may roots of unity as desired.
\end{proof}


\problem{13.6.6}
\begin{proof}
  The $2n$th roots of unity are the roots of $f(x) = x^{2n}-1 = (x^n-1)(x^n+1)$.
  But note that all the roots of $x^n - 1$ are $n$th roots of unity, so they aren't primitive $2n$th roots of unity, and hence aren't roots of $\Phi_{2n}$.
  Then the only possible primitive $2n$th roots of unity are the roots of $x^n + 1$. But note that since $n$ is odd we have $x^n + 1 = -((-x)^n - 1) = -\Phi_n(-x)$, so the roots of $x^n + 1$ are the negatives of the primitive $n$th roots of unity.
  But note that this implies that all these roots are primitive $2n$th roots, since for such root $\zeta$ any $k < n$ we have $-\zeta^{n+k} = -\zeta^n\zeta^k = \zeta^k$ which can't be $1$ by the fact $\zeta$ is a primitive $n$th root of unity.
  Therefore each of the roots of $x^n + 1$ are primitive $2n$th roots of unity, and so altogether we conclude that these are exactly the primitive $2n$th roots of unity.
  Finally then by the definition of cyclotomic polynomials we conclude that $\Phi_{2n}(x) = \Phi_n(-x)$, since the roots on both sides are exactly the primitive $2n$th roots of unit.
\end{proof}


\problem{13.6.9}
\begin{proof}
  Let $A$ be an $n \times n$ matrix over $\C$ for which $A^k = I$ for some integer $k \ge 1$. Then note that $A^k - I$, so its minimal polynomial divides $x^k - 1$, since otherwise we could construct an even smaller minimal polynomial via the division algortihm.
  But note that this polynomial is seperable, because its roots are the $k$th roots of unity.
  Hence any divisor of it is seperable, and in particular its minimal polynomial is seperable.
  Finally this implies that the eigenvalues of $A$ are distinct and so $A$ is diagonalizable.

  Then consider the matrix $A = \begin{pmatrix}1 & \alpha \\ 0 & 1\end{pmatrix}$ where $\alpha$ is an element of a field of characteristic $p$.
  We have that
  \begin{equation}
    \begin{pmatrix}
      1 & \alpha \\ 0 & 1
    \end{pmatrix}
    \begin{pmatrix}
      1 & k\alpha \\ 0 & 1
    \end{pmatrix} =
    \begin{pmatrix}
      1 & k\alpha + \alpha \\ 0 & 1
    \end{pmatrix} =
    \begin{pmatrix}
      1 & (k+1)\alpha \\ 0 & 1
    \end{pmatrix}
  \end{equation}
  Therefore by induction we conclude that $A^k = \begin{pmatrix}1 & k\alpha \\ 0 & 1\end{pmatrix}$. But then we note that since the field has characteristic $p$ we get $p\alpha = 0$, so that $A^p = I$ as desired.
  Furthermore we have that $(A-I)^2 = 0$, so the minimal polynomial for $A$ must divide $(x-1)^2$.
  But we also see $A-I \ne 0$, so the minimal polynomial can't be $x-1$.
  Therefore the minimal polynomial must be $(x-1)^2$, which is inseperable.
  Hence $A$ has repeated eigenvalues, and it can only be diagonalizable if it is equal to $\lambda I$, which clearly isn't the case. So we conclude $A$ is not diagonalizable.
\end{proof}


\problem{14.1.4}
\begin{proof}
  If there were an isomorphsim $\varphi: \Q(\sqrt{2}) \rightarrow \Q(\sqrt{3})$ then the image $\alpha = \varphi(\sqrt{2})$ would satisfy $\alpha^2 = \varphi(2) = \varphi(1) + \varphi(1) = 1 + 1 = 2$.
  Writing $\alpha$ in terms of the basis $\{1, \sqrt{3}\}$ for $\Q(\sqrt{3})$, we get that there exists $a,b \in \Q$ with $\alpha^2 = (a + b\sqrt{3})^2 = 2$. Expanding the LHS we get $a^2 + 3b^2 + 2ab\sqrt{3} = 2$. Equating the coefficients of $\sqrt{3}$ on both sides gives $2ab = 0$, so either $a = 0$ or $b = 0$.

  In the first case were $a = 0$, we get $3b^2 = 2$ and so $b^2 = 2/3$.
  Then note that when the LHS is written as a reduced fraction, the numerator and denominator will contain an even number of each prime factor, while on the RHS this isn't true, so we have a contradiction.
  On the other hand when $b = 0$ we get $a^2 = 2$, which contradicts the fact that $a \in \Q$. So in either case we get a contradiction and we conclude the two fields must not be isomorphic.
\end{proof}


\problem{14.2.4}
\begin{proof}
  Let $p$ be a prime, and let $f(x) = x^p - 2$.
  Then the roots of $f$ are of the form $\zeta_p^k\omega$, where $\zeta_p$ is a $p$th root of unity and $\omega = \sqrt[p]{2}$ is any fixed $p$th root of $2$.
  Note that we can just consider the case with $0 \le k < p$, since the powers of $\zeta_p$ form a cyclic group of order $p$.
  Then let $\sigma$ be an automorphism in the Galois group, and consider how it permutes these roots.
  By the definition of a field automorphism we have $\sigma(\zeta_p^k\omega) = \sigma(\zeta_p)^k\sigma(\omega)$, so it suffices to consider how $\sigma$ acts on $\zeta_p$ and $\omega$.
  
  Let $\sigma_{i,j}$ denote the automorphisms such that $\zeta_p$ goes to some power $\zeta_p^i$ with $1 \le i < p$ and $\omega$ goes to $\omega\zeta_p^j$ with $0 \le j < p$.
  Then $\sigma_{i,j}$ is a member of the Galois group since a root $\zeta_p^k\omega$ will map to $\zeta_p^{ki + j}\omega$ which is still a root.
  Note that this gives $p-1$ choices for $i$ and $p$ choices for $j$.
  Furthermore all these automorphisms are distinct since if the $i$ values differs than the image of  $\omega\zeta_p$ will differ and if $j$ values differ then the images of $\omega$ will differ.
  Hence this gives at least $p(p-1)$ distinct automorphisms in the Galois group.
  
  But the splitting field of $x^p - 2$ is also at most a degree $p(p-1)$ extension since we have $[\Q(\omega):\Q] = p$ and also that $[\Q(\zeta_p):\Q] = \deg(\Phi_p) = \varphi(p) = p-1$.
  Hence the number of automorphisms in the Galois group is exactly $p(p-1)$, and we conclude that the $p(p-1)$ distinct automorphisms in the previous paragraph are exactly the elements of the Galois group:
  \begin{equation*}
    G = \{ \sigma_{i,j} \}_{1 \le i < p;\ 0 \le j < p}
  \end{equation*}
\end{proof}


\problem{14.2.5}
\begin{proof}
  We define an isomorphism $h: G \rightarrow M$ from the group $G$ at the end of the last problem to group $M$ of matrices $\begin{pmatrix} a & b \\ 0 & 1 \end{pmatrix}$ with $a,b \in \F_p$, $a \ne 0$ as follows:
  \begin{equation*}
    \sigma_{i,j}
    \mapsto{h}
    \begin{pmatrix} i & j \\ 0 & 1 \end{pmatrix}
  \end{equation*}
  This map is well defined by the fact that the constraints on $i$ and $j$ in the definition of $G$ coincide exactly with the constraint that $a,b \in F_p$ with $a \ne 0$ (where we've identified $\F_p$ with the integers $0,...,p-1$ in the natural way).

  Then we show this is a group homomorphism.
  First, note that the identity element of the Galois group $G$ is the map $\parens*{\zeta_p^k\omega \mapsto \zeta_p^k\omega}$, which corresponds to $\sigma_{1,0}$.
  Then under $h$ this maps to the matrix $\begin{pmatrix} 1 & 0 \\ 0 & 1 \end{pmatrix}$, which is exactly the identity element of $M$, and so $h$ respects the identity elements.

  Then note that under circosition we have $\sigma_{i,j} \circ \sigma_{i',j'} = \sigma_{ii', ij' + j}$ since we have pointwise equality on each of the roots (Where we let $x_k = \zeta_p^k\omega$):
  \begin{equation*}
    \sigma_{i,j} \circ \sigma_{i',j'} (x_k)
    = \sigma_{i,j}(x_{i'k + j'})
    = x_{ii'k + ij' + j}
    = \sigma_{ii', ij' + j}(x_k)
  \end{equation*}
  Using this we get that:
  \begin{equation*}
    h(\sigma_{i,j} \circ \sigma_{i',j'} (x_k))
    = h(\sigma_{ii', ij' + j}(x_k))
    = \begin{pmatrix} ii' & ij' + j \\ 0 & 1 \end{pmatrix}
    = \begin{pmatrix} i & j \\ 0 & 1 \end{pmatrix}\begin{pmatrix} i' & j' \\ 0 & 1 \end{pmatrix}
    = h(\sigma_{i,j}) h( \sigma_{i',j'} (x_k))
  \end{equation*}
  And therefore $h$ respects the group operations, and we get that it
  is in fact a group homomorphism.

  Furthermore, we know that $h$ is a surjection, since for any values
  $a,b$ in the top row of the matrix, that will be the image of
  $\sigma_{a,b}$, which is an element of $G$ by the restriction that
  $a,b \in \F_p$ with $a \ne 0$.\.  But both sets are finite, so this
  implies it is bijective and so an isomorphism.  Finally we conclude
  that both groups are isomorphic by an explicit isomorphism.
\end{proof}


\problem{14.2.11}
\begin{proof}
  Let $f(x) \in \Z[x]$ be an irreducible quartic whose splitting field has Galois group $S_4$ over $\Q$.
  Let $\theta$ be a root of $f$ and let $K = \Q(\theta)$. 
\end{proof}


\problem{14.2.13}
\begin{proof}
  
\end{proof}


\problem{14.2.17}
\begin{proof}
  
\end{proof}


\problem{14.2.18}
\begin{proof}
  
\end{proof}


\problem{14.2.31}
\begin{proof}
  
\end{proof}


\end{document}