%&pdflatex
\documentclass[11pt]{article}

\usepackage{main-macros}

\usepackage{newpxmath}
\usepackage{newpxtext}
\usepackage[margin=0.75in]{geometry}

\newcommand{\disc}{\text{disc}} 

\title{8251 HW1}
\author{Devon Tuma}
\date{Fall 2020}

\begin{document}
\maketitle

\problem{1.16.}
Show that $(1 - \omega)(1 - \omega^2)\dots(1 - \omega^{p-1}) = p$.

\begin{proof}
  Setting $g(t) := (t - \omega)\dots(t - \omega^{p-1})$, we want to show that $g(1) = p$.
  From equation (1.2) we have that
  \begin{equation*}
    t^p - 1 = (t - 1)(t - \omega)\dots(t - \omega^{p-1}) = (t-1)g(x)
  \end{equation*}
  If we take formal derivatives of both sides, we get that $pt^{p-1} = g(x) + (t-1)g'(x)$, and evaluating this at $1$ we see that $p = g(1) + 0 * g'(1) = g(1)$, which is exactly what we wanted to show.
\end{proof}


\problem{1.21}
Show that every member of $\Q[\omega]$ is uniquely representable in the form $$a_0 + a_1\omega + \dots + a_{p-2}\omega^{p-2}, \ a_i \in \Q$$

\begin{proof}
  We need to show that $\{1, \omega, \dots, \omega^{p-2}\}$ is a $\Q$-basis for $\Q[\omega]$.
  
  To show that it is a spanning set, it suffices to show $\omega$ is a root of $f(t) = t^{p-1} + \dots + t + 1$, since then $\omega^{p-1} = -\omega^{p-2}-\dots-t-1$ and any power $\omega^m$ with $m \ge p - 1$ can be recursively reduced to a smaller power by substituting this equation.
  But note that we have $\omega f(\omega) = 1 + \omega^{p-1} + \dots + \omega^2 + \omega = f(\omega)$, and $\omega \ne 1$, so we must have $f(\omega) = 0$ as desired.

  Then to show the set is linearly independent, it suffices to show $f(t)$ is irreducible, since any nontrivial linear dependence among the $\{1, \omega, \dots, \omega^{p-2}\}$ defines a polynoial with $\omega$ as a root, of degree stricly less than $p-1$, which is then a nontrivial factor of $f(t)$.
  Furthermore, it suffices to show $f(t+1)$ is irreducible, since $g(t) \mid f(t)$ iff $f(t+1) \mid g(t+1)$. Then note that $(t-1)f(t) = t^p - 1$, so that $f(t+1) = ((t+1)^p - 1) / t)$.
  Applying the binomial thoerem, we see $f(t+1) = \sum_{n = 1}^p \binom{p}{n} t^{n-1}$.
  But then all coefficients but the first coefficient are divisible by $p$, and the last coefficient is exactly $p$ which isn't divisible by $p^2$, so we can apply Eisenstein's criterion to get irreducibility as desired.
\end{proof}

  
\problem{1.22}
Show that if $\alpha \in \Z[\omega]$ and $p \mid \alpha = \alpha_0 + \alpha_1\omega + \dots + \alpha_{p-2}\omega^{p-2}$, then all the $\alpha_i$ are divisible by $p$.

\begin{proof}
  If we have divisiblity, let $\beta$ be such that $p\beta = \alpha$.
  Since irreducibility in $\Q[x]$ implies irreducibility in $\Z[x]$, the previous exercise also implies a similar representation for elements of $\Z[x]$.
  In particular, $\beta$ has a representation of the form $\beta = \beta_0 + \beta_1\omega + \dots + \beta_{p-2}\omega^{p-2}$ with $\beta_i \in \Z$.
  But then by the linear independence of the powers of $\omega$, we must have $\alpha_i = p\beta_i$, and we conclude that $p | \alpha_i$ for all $i$.
\end{proof}

\problem{2.4}
Suppose $a_0, \dots, a_{n-1}$ are algebraic integers and $\alpha$ is a complex number satisfying $\alpha^n + a_{n-1}\alpha^{n-1} + \dots + a_0 = 0$. Show taht the ring $\Z[a_0, \dots, a_{n-1}, \alpha]$ has finitely generated additive group.

\begin{proof}
  We know that elements of $\Z[a_0, \dots, a_{n-1}, \alpha]$ are finite $\Z$-linear combinations of products of the form $a_0^{k_0} \dots a_{n-1}^{k_{n-1}} \alpha^k$, so it suffices to show that only many of these products are independent.
  By assumption, we have $\alpha^n = -a_{n-1}\alpha^{n-1} - \dots - a_0$, so we can certainly reduce any term with a power of $\alpha$ greater than $n$ to a linear combination of terms with powers of $\alpha$ strictly less than $n$ by recursively substituting this equation (each substitution strictly reduces the maximal power of $\alpha$ so this process terminates).

  But we also have that each $a_i$ is an algebraic integra, so there exists monic polynomials $p_i \in \Z[x]$ such that $p_i(a_i) = 0$ for all $i$.
  Each $p_i$ is monic, so is of the form $x^{deg(p_i)} + p'_i(x)$ where the trailing terms in $p'_i$ have degree strictly less than $deg(p_i)$.
  Hence we have $a_i^{deg(p_i)} = - p'_i(a_i)$, where each $p'_i(a_i)$ is a $\Z$-linear combination of powers of $a_i$, with each power of $a_i$ having degree less than $deg(p_i)$.
  So again we can recursively substitute these values, and the process will again eventually terminate since each substituion strictly reduces the maximal degree of a power of $a_i$.

  Hence after all of the substitutions, we can write any element as a linear combination of terms of the form $a_0^{k_0} \dots a_{n-1}^{k_{n-1}} \alpha^k$, with $k_i < deg(p_i)$ and $k < n$, and so this forms a finite generating set as desired.
  Then since $\alpha \Z[a_0, \dots, a_{n-1}, \alpha] \subset \Z[a_0, \dots, a_{n-1}, \alpha]$, we finally conclude that $\alpha$ is an algebraic integer as well.
\end{proof}


\problem{2.6}
Show that if $f$ and $g$ are polynomials over a field $K$ and $f^2 |\ g$ in $K[x]$ then $f |\ g'$.

\begin{proof}
  Since $f^2 |\ g$, we have $g = f^2h$ for some $h \in K[x]$.
  Taking formal derivatives, we have $$g' = 2ff'h+f^2h' = f(2f'h+fh')$$ and we see that $f |\ g'$ exactly as desired.
\end{proof}


\problem{2.10}
Show that if $m |\ r$ and $\varphi(r) \le \varphi(m)$ then $r = m$.

\begin{proof}
  Since $m |\ r$, we have $m \le r$, and so it suffices to show $r \le m$.

  Let $m = p_0^{e_0} p_1^{e_1} \dots p_n^{e_n}$ be the prime factorization of $m$.
  Then since $m |\ r$, the prime factorization of $r$ must be of the form $r = k * p_0^{e_0 + d_0} p_1^{e_1 + d_1} \dots p_n^{e_n + d_n}$ for some $k$ relatively prime to all of the $p_i$.
  Using the standard identities for $\varphi$, we then get:
  \begin{equation*}
    \varphi(m) = \prod_{0 \le i \le a} (p_i^{e_i} - p_i^{e_i - 1}) \le
    \varphi(k) \prod_{0 \le i \le a} (p_i^{e_i + d_i} - p_i^{e_i + d_i - 1})
    = \varphi(r)
  \end{equation*}
  And we already had $\varphi(m) \ge \varphi(r)$, so we have in fact $\varphi(m) = \varphi(r)$.
  But then looking at the above equation egain, we see that equality holds only if $d_i = 0$ for all $i$ and $\varphi(k) = 1$.
  $\varphi(k) = 1$ implies $k = 1$, and then if we plug these values back into the oringal prime factorizations we see that they are now identical, and we conclude $r = m$ as desired.
\end{proof}


\problem{2.25}
Show that for any algebraic number $\alpha$, there exists $m \in \Z$, $m \ne 0$, such that $m\alpha$ is an algebraic integer.

\begin{proof}
  Let $\alpha$ be algebraic, and $p \in \Q[x]$ a witness to this, so $p(\alpha) = 0$.
  Clearing denominators on the coefficients of $p$, we can write $p = kf$, for some $f \in \Z[x]$ and $k \in \Q$.
  Note that $f(\alpha) = p(\alpha)/k = 0$, so $\alpha$ is a root of $f$.
  Hence $0 = f(\alpha) = c_n\alpha^n + \dots + c_1\alpha + c_0$, where the coefficeients $c_i$ are all integers.
  We will show $c_n\alpha$ is integral ($c_n$ is non-zero since $p$, and hence $f$, is nontrivial).
  
  Multiplying the equation by $c_n^{n-1}$ we get:
  \begin{align*}
    0 = c_n^{n-1} f(\alpha)
    &= c_n^n\alpha^n + c_{n-1}c_n^{n-1}\alpha^{n-1} + c_{n-2}c_n^{n-1}\alpha^{n-2} + \dots + c_0c_n^{n-1} \\
    &= (c_n\alpha)^n + c_{n-1}(c_n\alpha)^{n-1} + c_{n-2}c_n(c_n\alpha)^{n-2} + \dots + c_0c_n^{n-1}
  \end{align*}
  And we see that $c_n\alpha$ is a root of the polynomial $x^n + c_{n-1}x^{n-1} + c_{n-1}c_nx^{n-2} + \dots + c_0 c_n^{n-1}$, which is monic and has integer coefficients.
  Hence $c_n\alpha$ is an algebraic integer, as desired.
\end{proof}


\problem{2.28}
Let $f(x) = x^3 + ax + b$, with $a,b \in \Z$ and assume $f$ is irreducible over $\Q$ with $\alpha$ a root of $f$.

\textbf{(a)} Show that $f'(\alpha) = -(2a\alpha + 3b)/\alpha$.

\begin{proof}
  We have $f'(x) = 3x^2 + a$, so $f'(\alpha) = 3\alpha^2 + a = (3\alpha^3 + a\alpha)/\alpha$.
  But since $\alpha$ is a root of $f$, we also know $\alpha^3 = -a\alpha - b$, and substituting this into the previous equation gives $f'(\alpha) = -(3a\alpha - 3b - a\alpha)/\alpha = -(2a\alpha + 3b)/\alpha$ as desired.
\end{proof}

\textbf{(b)} Show that $2a\alpha+3b$ is a root of $(\frac{x-3b}{2a})^3+a(\frac{x-3b}{2a})+b$, Use this to find $N_\Q^{\Q[\alpha]}(2a\alpha+3b)$.

\begin{proof}
  Note that the given equation is just $f(g(x))$ where $g(x):=(x-3b)/2a$.
  Furthermore, we have $g(2a\alpha+3b)=2a\alpha/2a=\alpha$, so we have $f(g(2a\alpha+3b)) = f(\alpha) = 0$ by definition of $\alpha$.
  Hence $2a\alpha+3b$ is a root of $f \circ g$ as desired.
\end{proof}

  Then this must be the minimal polynomial for $2a\alpha+3b$ since we assumed $f$ was irreducible, so to find $N_\Q^{\Q[\alpha]}(2a\alpha+3b)$ we just need to find the product of all its roots (since these roots are exactly the conjugates).
  But note that after normalizing by the leading coefficient to make the polynomial monic, this is just the constant term of the resulting polynomial.
  The leading coefficient is $1/8a^3$, and the constant term is $-27b^3/8a^3-3b/2+b$, so we conclude that $N_\Q^{\Q[\alpha]}(2a\alpha+3b) = -27b^3-12ba^3 + 8ba^3 = -27b^3-4ba^3$.

\textbf{(c)} Show that $\disc(\alpha) = -(4a^3+27b^2)$

\begin{proof}
  By theorem 8, we have $\disc(\alpha) = N(f'(\alpha))$.
  Then by part (a) we know $f'(\alpha) = -(2a\alpha + 3b)/\alpha$, so using multiplicativity of the norm, we have $\disc(\alpha) = -N(2a\alpha + 3b)/N(\alpha)$.
  Furthermore, we know $N(\alpha) = b$ since $f$ is irreducible and monic, and plugging in this and the result of (b) gives $\disc(\alpha) = (-27b^3-4ba^3)/b = -(4a^3 + 27b^2)$ as desired.
\end{proof}

\textbf{(d)} Suppose $\alpha^3 = \alpha + 1$. Prove that $\{1,\alpha,\alpha^2\}$ is an integral basis for $\mathbb{A} \cap \Q[\alpha]$.

\begin{proof}
  If $\alpha^3 = \alpha + 1$ then $\alpha$ is a root of $x^3 - x - 1$. This is irrudicible since it has no roots, so we can apply (c) to get $\disc(\alpha) = -23$.
  Since this is squarefree, we can apply 27.e to get that $\{1,\alpha,\alpha^2\}$ is an integral basis.
  If instead $\alpha^3 + \alpha = 1$, then $\alpha$ is instead a root of $x^3 + x - 1$, so $\disc(\alpha) = -31$, which is again squarefree so we have the same integral basis.
\end{proof}


\problem{2.37}
Let $\alpha$ be an algebraic number of degree $n$ over $\Q$ and lef $f$ and $g$ be two polynomials over $\Q$, each of degree < $n$, such that $f(\alpha) = g(\alpha)$. Show that $f = g$.

\begin{proof}
  Since $f(\alpha) = g(\alpha)$, we have $(f - g)(\alpha) = 0$.
  Furthermore, we have $deg(f-g) \le max\{deg(f),\ deg(g)\} < n$.
  But $\alpha$ is of degree $n$, so it can't be a solution to a non-trivial polynomial of degree strictly less than $n$.
  Therefore we conclude $f - g = 0$ and so finally $f = g$ as desired.
\end{proof}


\problem{3.2}
Show that any finite integral domain is a field

\begin{proof}
  Let $R$ be a finite integral domain and let $x \in R$ be non-zero.
  Consider the sequence of elements $x^1, x^2, x^3, \dots$.
  Since $R$ is finite, this must eventually repeat, so eventually we have $x^n = x^{n + k}$ for some $k > 0$.
  Hence $x^n * 1 = x^n * x^k$, and since $R$ is an integral domain we can cancel the $x^n$ to get $1 = x^k$ ($x^n$ is non-zero since $x$ is non-zero).
  Finally then $x * x^{k-1} = 1$, and since each element has an inverse we conclude $R$ is a field.
\end{proof}


\problem{3.3}
Let $G$ be a free abelian group of rank $n$, with additive notation. Show that for any $m \in \Z$, $G/mG$ is the direct sum of $n$ cyclic groups of order $m$

\begin{proof}
  Since $G$ is free abelian of rank $n$, we have $G \cong \Z^n$, and so $G/mG \cong \Z^n / m \Z^n$, and we just need to show that $\Z^n / m \Z^n \cong (\Z/m\Z)^n$.
  We have a surjection $\varphi$ from $\Z^n$ to $(\Z/m\Z)^n$ defined by the sum of the surjective projection maps in each coordinate.
  By the first isomorphism theorem then $\Z^n / \ker \varphi \cong (\Z/m\Z)^n$, and so it suffices to check $\ker \varphi = m \Z^n$.
  But note that $\varphi(x) = 0$, iff every entry in $x$ must be divisible by $m$, and so iff $x$ is an element of $m\Z^n$, as desired.
\end{proof}


\problem{3.7}
Show that if $I$ and $J$ are ideals in a commutative ring $R$ such that $1 \in I + J$, then $1 \in I^m + J^n$ for all $m,n$.

\begin{proof}
  Since $1 \in I + J$, we must have an expression $1 = \alpha + \beta$, with $\alpha \in I$ and $\beta \in J$.
  Then for any fixed $m$ and $n$, we have:
  \begin{align*}
    1 = 1^{m + n}
    &= (\alpha + \beta)^{m + n} \\
    &= \alpha^{m+n} + \dots + c_{m+1}\alpha^{m+1}\beta^{n-1} +  c_{m}\alpha^m\beta^n + \dots + \beta^{m+n} \\
    &= \alpha^m (\alpha^n + \dots + c_{m+1}\alpha^{1}\beta^{n-1}) + \beta^n (c_{m}\alpha^m + \dots + \beta^{m})    
  \end{align*}
  Where the $c_i = \binom{n+m}{i}$. 
  Therefore we see that $1$ is a linear combination of $\alpha^m \in I^m$ and $\beta^n \in I^n$, and so $1 \in I^m + J^n$ as desired.
\end{proof}


\problem{3.9}
Let $K$ and $L$ be number fields, $K \subset L$, $R := \mathbb{A} \cap K$, $S := \mathbb{A} \cap L$
\textbf{(a)} Let $I$ and $J$ be ideals of $R$, and suppose $IS \mid JS$. Show $I \mid J$.

\begin{proof}
  Let $I = P_1P_2 \dots P_n$ be the prime factorization of $I$ in $R$ (The $P_i$ need not be distinct), and similarily let $J = Q_1Q_2 \dots Q_m$ be the prime factorization of $J$ in $R$.
  We need to show that each $P_i$ corresponds to some $Q_k$.
  When we move to $S$, each factor $P_iS$ will split into a further prime ideals, which must all appear in the prime factorization of $JS$ since $IS \mid JS$.
  But by uniqueness of prime factorization, the factors of $P_iS$ must all come from the same $Q_jS$ factor of $JS$, since otherwise we would get a distinct representation of $J$ as a product of prime ideals, even though we know this representation is unique.
  Therefore since each $P_i$ has a corresponding $Q_j$ with identical factors in $S$, we conclude that $I \mid J$.  
\end{proof}

\textbf{(b)} Show that for any ideal $I$ in $R$, we have $I = IS \cap R$.

\begin{proof}
  Certainly we have $I \subseteq IS \cap R$, since for any $x \in I$ we have $x = x*1 \in IS$ and $x \in R$.
  Then for the other inclusion $IS \cap R \subseteq I$, it suffices by theorem 19 to show $I \mid IS \cap R$, and so by part (a) it suffices to show $IS \mid (IS \cap R)S$.
  But then by theorem 19 again, it suffices to show $IS \cap R \mid IS \cap R$, which is trivially true, so we're done.
\end{proof}

\textbf{(c)} Characterize those ideals $I$ of $S$ such that $I = (I \cap R)S$.

\begin{proof}
  First, we always have $(I \cap R)S \subseteq I$, since by theorem 19 this is equivalent to $I \cap R = I \cap R$, which is clearly true.
  Therefore we just need to characterize the ideals such that $I \subseteq (I \cap R)S$.
  This is exactly the ideals of $S$ where every element of $I$ is a finite $S$-linear combination of elements of $R$.
  This certainly includes the ideals that are images of ideals in $R$, so includes all ideals of the form $IS$ for $I$ an ideal of $R$.
  But it also can't include any more ideals, since it always suffices to find just a finite set of generators, because for any infinite generating set we can find a basis by viewing the ideal as an $R$-submodule.
  Therefore this is an exact characterization of these ideals.
\end{proof}


\problem{3.16}
Let $K$ and $L$ be number fields, $K \subset L$, $R = \mathbb{A} \cap K$, $S = \mathbb{A} \cap L$. Denote by $G(R)$ and $G(S)$ the ideal class groups of $R$ and $S$, respectively.

\textbf{(a)} Show that there is a homomorphism $G(S) \rightarrow G(R)$ defined by taking any $I$ in a given class $C$ and sending $C$ to the class containing $N_K^L(I)$.

\begin{proof}
  First, $N_K^L$ is multiplicative and sends principle ideals to principle ideals, by problem 15.
  Therefore we need to only check that this is a well defined function on the quotient.
  So let $I$ and $J$ be members of the same ideal class, so $\alpha I = \beta J$ for some $\alpha, \beta$.
  Applying norms to both sides we have $N_K^L(\alpha I) = N_K^L(\beta J)$.
  But then by the properties of $N_K^L$ established in problem 15, we have $(N_K^L(\alpha) N_K^L(I) = (N_K^L(\beta)) N_K^L(J)$, so the images of $I$ and $J$ lie in the same ideal class as well, and the function is indeed well defined.
\end{proof}

\textbf{(b)} Let $Q$ be a prime of $S$ lying over a prime $P$ or $R$. Let $d_Q$ denote the order of the class containing $Q$ in $G(S)$, $d_P$ the order of the class containing $P$ in $G(R)$. Prove that $d_P \mid d_Q f(Q|P)$.

\begin{proof}
  First, by definition of the ideal norm we have $N_K^L(Q) = P^{f(Q|P)}$.
  We need to show that $d_P \mid d_Q f(Q|P)$, and it suffices to show $d_P/d_Q = k f(Q|P)$ for some $k$.
  But by part (a) the ideals in a given ideal class in $K$ in have a density of at most $f(Q|P)$ in the corresponding ideal class in $L$, so this holds and we're done.
\end{proof}


\problem{3.25}
Let $R$ be a commutative ring, $I$ an ideal; for each $f \in R[x]$, let $\bar f$ denote the image of $f$ under the homomorphism $R \rightarrow (R/I)[x]$.

\textbf{(a)} Show that $\bar f = \bar g$ iff $f - g \in I[x]$.

\begin{proof}
  Let $f_i$ and $g_i$ denote the $i$th coefficient of the respective polynomial.
  Then polynomials are determined uniquely by their coefficients, so $\bar f = \bar g$ iff $\bar f_i = \bar g_i$ for all $i$, where the quotient is now taken in just $R/I$.
  But then equality in the quotient implies that $\bar f_i = \bar g_i$ if and only if $f_i - g_i \in I$ for all $i$, which then implies that $f - g \in I[x]$, since all of its coefficients are in $I$, so we are done.
\end{proof}

\textbf{(b)} Show that $\bar g \mid \bar f$ iff $f \in (I,g)$.

\begin{proof}
  We have $\bar g \mid \bar f$ iff $\bar{gh} = \bar g \bar h = \bar f$ for some $\bar h \in (R/I)[x]$.
  By definion of equality in the quotient, this happens iff $f - gh \in I[x]$, or equivalently iff $f = gh + k$ for some $k \in I[x]$.
  But then altogether this happens iff $f \in (g, I)$ since elements of $(g,I)$ are linear combinations of $g$ and an element of $I$, which is exactly what we wanted to show.
\end{proof}

\textbf{(c)} Show that $R[x]/(I,g)$ is isomorphic to $(R/I)[x]/(\bar g)$.

\begin{proof}
  Define $\varphi : R[x] \rightarrow (R/I)[x]/(\bar g)$ to be the composition of the two corresponding projection functions to the quotient.
  This is a surjective ring homomorphism, since it is the composition of projections which are both surjective.
  So by the first isomorphism theorem, we have $R[x]/(\ker \varphi) \cong (R/I)[x]/(\bar g)$, and so it suffices for us to show $\ker \varphi = (I,g)$.

  Then we have $f \in \ker \varphi$ iff $\bar f \in (\bar g)$, and therefore iff $\bar g \mid \bar f$. But then by (b) this happens iff $f \in (I,g)$.
  Hence we conclude $\ker \varphi = (I, g)$, which is what we needed to show.
\end{proof}


\end{document}
